---
title: 'Practical Machine Learning: Prediction Assignment Writeup'
author: "Michal Smolira"
date: "27.10.2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.



## Data Preparation

FIrst we load the data from above urls . At first glance we see that there is many fields with blank space , and few with value ,“#DIV/0” so both we change to NA values

```{r}
pml_training<-read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"),na.strings=c("","NA","#DIV/0"))

pml_testing<-read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"),na.strings=c("","NA","#DIV/0"))


```

Calculate percent of NA values for each column

```{r}
Columquality<-colSums(is.na(pml_training))/nrow(pml_training)

```

Remove columns with more than 5% NA Values

```{r}
pml_training<-pml_training[Columquality<0.05]
pml_testing<-pml_testing[Columquality<0.05]

```

Remove first 7 columns which describe the data rows


```{r}
pml_training<-pml_training[,-c(1:7)]
pml_testing<-pml_testing[,-c(1:7)]

```



```{r}
dim(pml_training)


```
The final training dataset has 19622 rows and 53 variables. We have 52 predictors



## Cross- Validation

The training dataset will be divide to two data set 
* Training (70% of original pml_training)
* Validation (30% of original pml_training)


```{r}

set.seed(1109)

library(caret)

inTrain <- createDataPartition(y=pml_training$classe,p=0.7,list=FALSE)

training<- pml_training[inTrain,]

validation<-pml_training[-inTrain,]


```



## Model Chose


### Because predicting value is a factor we start building models from Decision Trees




```{r}

library(rpart)

ModelTrees<- rpart(classe~.,data = training,method = "class")

library(rattle)

fancyRpartPlot(ModelTrees)


predictTrees<-predict(ModelTrees,newdata = validation,type = "class")


confusionMatrix(predictTrees,validation$classe)

```
**Accuracy for decision Trees is 0.7397**




###The second model use Random forest method 

```{r}


library(randomForest)

RandomForest <- randomForest(classe ~ ., data = training)

predictRF<-predict(RandomForest,newdata = validation)


confusionMatrix(predictRF,validation$classe)

```


**Accuracy for decision Trees is 0.9949**


## Summary

Random Forest method is much better then Decision Three **0.9949** vs **0.7397**.We can even assume that we achive 100% accuracy for pml_testing dataset.



```{r}

finalcheck <- predict(RandomForest, newdata = pml_testing, type="class")
finalcheck

```



